Xinhua, Diplomacy, and Chinese Foreign Aid

### 2022-03-29 04_paper/draft1.Rmd
more updates from MW

---
title: "Xinhua, Diplomacy, and Chinese Foreign Aid"
author: ["Lucie Lu","Miles D. Williams"]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: yes
classoption: fleqn
fontfamily: mathpazo
fontsize: 12pt
bibliography: ref.bib
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \setlength{\mathindent}{15pt}
  - \setlength\parindent{24pt}
  - \setlength{\parskip}{0.0pt plus 1.0pt}
  - \doublespacing
  - \newtheorem{theorem}{Theorem}[section]
  - \newtheorem{lemma}[theorem]{Lemma}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{corollary}[theorem]{Corollary}
  - \newenvironment{proof}[1][Proof]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries Proof}]}{\end{trivlist}}
  - \newenvironment{definition}[1][Definition]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{example}[1][Example]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{remark}[1][Remark]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{!t}
abstract: "Absent formal avenues of transparency, how do China's efforts at gaining international recognition and promoting good relations with aid recipients play out in Beijings foreign aid giving? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. These measures capture two distinct but related aspects of China’s foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. To the extent that China’s development finance complements these goals, we expect greater coverage of a developing country in a Chinese media outlet directed to foreign readers and greater bilateral diplomatic activity to correlate with greater Chinese foreign aid giving. To test this expectation we merge AidData’s Chinese development finance dataset with two newly available datasets: (1) AidData’s compiled yearly counts of bilateral official diplomatic visits from China and (2) millions of English edition Xinhua news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research. We find that greater media coverage of developing countries targeted at a foreign audience via the English version of Xinhua and the number of diplomatic visits to a developing country predict greater receipt of Chinese aid. Our results support the view that Chinese aid allocation patterns map to China’s broader diplomacy and legitimacy seeking objectives on the world stage."
---

@@

For our empirical strategy, we estimate models of the form
$$\text{ihs(aid}_{rt}) = \tau_t + \beta_1\text{ihs(coverage}_{rt}) + \beta_2\text{ihs(visits}_{rt}) + X_{rt}^\top \gamma + \epsilon_{rt}. \tag{1}$$
In the above, subscript $r$ indexes the aid recipient and $t$ the year. The vector $X_{rt}$ is the set of recipient and bilateral controls, and $\epsilon_{rt}$ is a zero-centered stochastic term.

@@

In the event that the process that determines whether countries receive aid from China differs from the process that determines the amount of aid given, we further separate the estimation into selection and level equations as outlined by Cragg (year?). The selection equation is estimated using mixed effects logit with random recipient intercepts and the level equation is estimated using OLS with cluster-robust standard errors.

Finally, because within-year coverage and visits may have greatest relevance for predicting the response, using the lag of each variable in model estimation may under-estimate the complementarities between coverage and visits with aid. However, estimating the models using same-year coverage and visits poses a challenge due to likely endogeneity between them and the response. Unable to find a satisfactory set of instruments

@@

```{r, results='asis'}
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"), "/03_analysis/tab1.R")
)
reg.tab1 <- reg.tab
cat(reg.tab1)
```

@@

>>>>

---
title: "Xinhua, Diplomacy, and Chinese Foreign Aid"
author: ["Lucie Lu","Miles D. Williams"]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: yes
classoption: fleqn
fontfamily: mathpazo
fontsize: 12pt
bibliography: ref.bib
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \setlength{\mathindent}{15pt}
  - \setlength\parindent{24pt}
  - \setlength{\parskip}{0.0pt plus 1.0pt}
  - \doublespacing
  - \newtheorem{theorem}{Theorem}[section]
  - \newtheorem{lemma}[theorem]{Lemma}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{corollary}[theorem]{Corollary}
  - \newenvironment{proof}[1][Proof]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries Proof}]}{\end{trivlist}}
  - \newenvironment{definition}[1][Definition]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{example}[1][Example]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{remark}[1][Remark]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{!t}
  - \usepackage{caption}\captionsetup[table]{skip=5pt}
  - \usepackage{dcolumn}
  - \usepackage{caption}\DeclareCaptionLabelFormat{AppendixTables}{A.#2}

abstract: "Absent formal avenues of transparency, how do China's efforts at gaining international recognition and promoting good relations with aid recipients play out in Beijings foreign aid giving? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. These measures capture two distinct but related aspects of China’s foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. To the extent that China’s development finance complements these goals, we expect greater coverage of a developing country in a Chinese media outlet directed to foreign readers and greater bilateral diplomatic activity to correlate with greater Chinese foreign aid giving. To test this expectation we merge AidData’s Chinese development finance dataset with two newly available datasets: (1) AidData’s compiled yearly counts of bilateral official diplomatic visits from China and (2) millions of English edition Xinhua news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research. We find that greater media coverage of developing countries targeted at a foreign audience via the English version of Xinhua and the number of diplomatic visits to a developing country predict greater receipt of Chinese aid. Our results support the view that Chinese aid allocation patterns map to China’s broader diplomacy and legitimacy seeking objectives on the world stage."
---

@@

For our empirical strategy, we estimate models of the form
$$\text{ihs(aid}_{rt}) = \tau_t + \beta_1\text{ihs(coverage}_{rt}) + \beta_2\text{ihs(visits}_{rt}) + X_{rt}^\top \gamma + \epsilon_{rt}. \tag{1}$$
In the above, subscript $r$ indexes the aid recipient and $t$ the year. The vector of parameters $\tau_t$ denotes fixed year intercepts, the vector $X_{rt}$ is the set of recipient and bilateral controls, and $\epsilon_{rt}$ is a zero-centered stochastic term.

@@

In the event that the process that determines whether countries receive aid from China differs from the process that determines the amount of aid given, we further separate the estimation into selection and level equations as outlined by @cragg1971. The selection equation is estimated using mixed effects logit with random recipient intercepts and the level equation is estimated using OLS with cluster-robust standard errors.

Finally, because within-year coverage and visits may have greatest relevance for predicting the response, using the lag of each variable in model estimation may under-estimate the complementarities between coverage and visits with aid. However, estimating the models using same-year coverage and visits poses a challenge due to likely endogeneity between them and the response. Unable to find a satisfactory set of instruments, we take two different approaches. For one, we leverage the variation in the one year lag of each predictor to instrument for in-year values of each. This approach has been applied elsewhere in the literature when dealing with time-series data [see @braumoeller2008]. For the second approach, we use the internal instruments approach outlined by @lewbel2012, which leverages heteroskedasticity in the data to synthesize instrumental variables.

We present results from each of these approaches in the following analysis. Further, in the Appendix we break down the sample by geographic region in the event that different dynamics govern decisions within different sets of aid recipients.

@@

```{r}
library(tidyverse)
library(kableExtra)
dt <- read_csv(
  paste0(str_remove(getwd(), "/04_paper"), 
         "/01_data/final_data/final_data_imputed.csv")
)
```

Results are shown for `r scales::comma(nrow(dt %>% filter(year!=2000)))` unique recipient-year observations from 2001 to 2014. Multiple imputation for missing values in covariates was done via chaining random forests combined with predictive mean matching, which ensures imputed values do not differ from values already present in the data (such as in the case of a binary variable) and that the conditional distributions have realistic variances.^[For multiple imputation we used the `missRanger` R package. See @stekhovenBuhlmann2011 for more on the method and its robustness relative to other approaches.] Summary statistics are shown the Appendix.

```{r, include=FALSE}
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"), "/03_analysis/tab1.R")
)
reg.tab1 <- reg.tab
cat(reg.tab1)
```

\begin{table}
\caption{Model Estimates}
\begin{center}
\begin{tabular}{l c c c c c c}
\hline
 & OLS & ML Tobit & Selection & Level & IV Lag & IV Lewbel \\
\hline
Coverage (lag)              & $0.63^{*}$      & $1.29^{**}$    & $0.18^{***}$  & $0.11$          &                 &                 \\
                            & $(0.28)$        & $(0.43)$       & $(0.05)$      & $(0.14)$        &                 &                 \\
Coverage (in-year)          &                 &                &               &                 & $0.67^{*}$      & $0.51^{\cdot}$  \\
                            &                 &                &               &                 & $(0.31)$        & $(0.29)$        \\
Diplomatic Visits (lag)     & $0.93^{**}$     & $1.54^{\cdot}$ & $0.21^{**}$   & $0.36^{*}$      &                 &                 \\
                            & $(0.34)$        & $(0.83)$       & $(0.07)$      & $(0.15)$        &                 &                 \\
Diplomatic Visits (in-year) &                 &                &               &                 & $1.08^{**}$     & $1.38^{**}$     \\
                            &                 &                &               &                 & $(0.41)$        & $(0.45)$        \\
Freedom House               & $-0.36$         & $-0.58$        & $-0.08^{*}$   & $-0.20^{*}$     & $-0.36$         & $-0.39$         \\
                            & $(0.28)$        & $(0.40)$       & $(0.04)$      & $(0.09)$        & $(0.28)$        & $(0.29)$        \\
GDP                         & $-2.71^{***}$   & $-4.83^{***}$  & $-0.77^{***}$ & $0.01$          & $-2.73^{***}$   & $-2.68^{***}$   \\
                            & $(0.36)$        & $(0.48)$       & $(0.06)$      & $(0.17)$        & $(0.36)$        & $(0.34)$        \\
Population                  & $2.52^{***}$    & $5.26^{***}$   & $0.69^{***}$  & $0.27$          & $2.51^{***}$    & $2.56^{***}$    \\
                            & $(0.35)$        & $(0.76)$       & $(0.06)$      & $(0.17)$        & $(0.35)$        & $(0.36)$        \\
Unemployment                & $0.18$          & $0.84$         & $0.13$        & $-0.05$         & $0.19$          & $0.20$          \\
                            & $(0.47)$        & $(0.95)$       & $(0.08)$      & $(0.19)$        & $(0.47)$        & $(0.47)$        \\
Disaster Deaths             & $0.18$          & $0.14$         & $0.04$        & $0.02$          & $0.18$          & $0.17$          \\
                            & $(0.17)$        & $(0.27)$       & $(0.03)$      & $(0.05)$        & $(0.17)$        & $(0.16)$        \\
Civil War                   & $-1.62^{\cdot}$ & $-2.77^{*}$    & $-0.38^{*}$   & $-0.75^{\cdot}$ & $-1.62^{\cdot}$ & $-1.57^{\cdot}$ \\
                            & $(0.91)$        & $(1.25)$       & $(0.15)$      & $(0.39)$        & $(0.92)$        & $(0.91)$        \\
Trade                       & $0.14$          & $-0.18$        & $0.04$        & $0.03$          & $0.14$          & $0.14$          \\
                            & $(0.15)$        & $(0.18)$       & $(0.03)$      & $(0.05)$        & $(0.15)$        & $(0.15)$        \\
Alliance                    & $0.36$          & $2.00$         & $0.29$        & $-0.64$         & $0.07$          & $-0.32$         \\
                            & $(0.94)$        & $(2.75)$       & $(0.22)$      & $(0.47)$        & $(0.94)$        & $(1.00)$        \\
Distance                    & $1.84^{*}$      & $4.12^{\cdot}$ & $0.49^{***}$  & $0.55$          & $1.94^{*}$      & $2.21^{*}$      \\
                            & $(0.83)$        & $(2.34)$       & $(0.15)$      & $(0.37)$        & $(0.87)$        & $(0.87)$        \\
\hline
Adj. R$^2$                  & $0.19$          &              &             & $0.11$          & $0.20$          & $0.20$          \\
Num. obs.                   & $1932$          & $1932$         & $1932$        & $982$           & $1932$          & $1932$          \\
Deviance explained          &               &              & $0.15$        &               &               &               \\
\hline
\multicolumn{7}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$; $^{\cdot}p<0.1$}}
\end{tabular}
\label{table:coefficients}
\end{center}
\end{table}

Table 1 reports the regression coefficients for equation 1 using each of the estimation techniques outlined in the previous section. Estimates for the controls comport with prior research, save that the coefficients for democracy using the Freedom House measure reach conventional levels of statistical significance in the selection and level estimates. Contra @dreherEtAl2018, these estimates suggest China disproportionately targets its development finance to authoritarian regimes. While this result is not statistically significant for the alternative estimators, the coefficient remains consistently negative across all estimation techniques.

Consistent with other studies, China prioritizes recipients with smaller economies and larger populations. Depending on the estimation strategy, greater distance is related to *more* aid, which is a finding that runs contrary to patterns observed for DAC donors. This result likely is driven by China's prioritization of aid giving in Sub-Saharan Africa. Finally, China seems to have a bias against giving aid in the context of ongoing conflict. Across estimation strategies, the coefficient on civil war is negative and marginally significant ($p < 0.1$) in four of the models and statistically significant ($p < 0.05$) in two.

China does not appear to discriminate on the basis of other measures of need, such as unemployment and natural disaster deaths. Further, trade and alliances, measures of material and strategic importance to Beijing, show no significant relationship with aid.

Turning to the estimates of interest, across specifications the coefficients on media coverage in *Xinhua* and the number of diplomatic visits run in the expected direction. Expressed as elasticities, a majority of the estimates suggest that a one percent greater yearly instances of coverage are associated with more than half a percent to over one percent greater aid received by an aid recipient relative to another aid recipient. Further, one percent greater diplomatic visits is related to almost one percent to one and a half percent greater aid to that recipient relative to another.

@@

# Appendix

\captionsetup{labelformat=AppendixTables}
\setcounter{table}{0}

```{r}
dt %>%
  mutate(
    aid = aid / 1000000,
    count = count / 1000,
    trade = trade / 1000000000,
    gdp = gdp / 1000000,
    pop = pop / 1000000
  ) %>%
  summarize(
    across(
      c(aid, count, govt_visits, pmm_fh, gdp:unemp, disaster:dist, trade, atopally),
      list(Mean = mean, Median = median, SD = sd, Min. = min, Max. = max),
      na.rm=T
    ) 
  ) %>%
  pivot_longer(cols = everything()) %>%
  separate(
    col = name,
    into = c(" ", "stat"),
    sep = "_(?=[^_]+$)"
  ) %>%
  mutate(
    value = scales::comma(round(value, 2))
  ) %>%
  pivot_wider(
    values_from = value,
    names_from = stat
  ) %>%
  mutate(
    " " = c(
      "Aid (millions)",
      "Article Mentions (thousands)",
      "Government Visits",
      "Freedom House",
      "GDP (in 100 millions)",
      "Population (in millions)",
      "Unemployment Rate",
      "Disaster Deaths",
      "Civil War",
      "Distance (in kilometers)",
      "Trade (in millions)",
      "Allies"
    )
  ) %>%
  kbl(
    caption = "Summary Statistics (2000-2014)",
    booktabs = T,
    linesep = ""
  )
```
